AAAI Press Formatting Instructions \\for Authors Using \LaTeX{} --- A Guide

### 2021-09-15 publications/iaai21/dora_outlier_detection_iaai.tex
Add isolation forest description. #42

\paragraph{Data types.} We chose to initially implement data loaders for four
data types that are commonly used by the machine learning and domain science
communities: time series, feature vectors, images (grayscale or RGB), and 
$N$-band rasters. $N$-band rasters are essentially images in which every pixel
is associated with a location defined by a geographic coordinate reference 
system (e.g., latitude/longitude in degrees). Most satellite data are 
distributed as rasters; common formats include GeoTIFF, NetCDF, and 
\todo{HDF}. A data loader for each data type locates the data by the path(s)
defined in the configuration file and loads samples into a dictionary of numpy
 arrays indexed by the sample id. This \texttt{data\_dict} is then passed to each of the ranking algorithms specified in 
 the configuration file.

@@

\subparagraph{Sparsity.}
\todo{Describe isolation forest: Umaa.}

\subparagraph{Likelihood.}
\todo{Describe PAE: Hannah or Bryce}
The negative sampling algorithm is implemented by converting the unsupervised 
outlier ranking problem into a semi-supervised 
problem~\citep{sipple:neg-sampling20}. Negative (anomalous) 
examples are created by sampling from an expanded space defined by the minimum 
and maximum values of each dimension of the positive (normal) examples. The 
negative and positive examples are then used to train a random forest 
classifier. We use the posterior probabilities of the random forest classifier 
as outlier scores, which means that the observations with higher posterior 
probabilities are more likely to be outliers.

>>>>

\paragraph{Data types.} We chose to initially implement data loaders for four
data types that are commonly used by the machine learning and domain science
communities: time series, feature vectors, images (grayscale or RGB), and 
$N$-band rasters. $N$-band rasters are essentially images in which every pixel
is associated with a location defined by a geographic coordinate reference 
system (e.g., latitude/longitude in degrees). Most satellite data are 
distributed as rasters; common formats include GeoTIFF, NetCDF, and 
HDF. A data loader for each data type locates the data by the path(s)
defined in the configuration file and loads samples into a dictionary of numpy
 arrays indexed by the sample id. This \texttt{data\_dict} is then passed to each of the ranking algorithms specified in 
 the configuration file.

@@

\subparagraph{Sparsity.}
Sparsity-based methods score outliers based on how isolated or sparse they
are in the feature space. Isolation forests are a common sparsity-based method
that constructs many random binary trees from a 
dataset~\cite{liu2008isolation}. The outlier score for
a sample is quantified as the average distance from the root to the item’s 
leaf. Shorter distances are indicative of outliers because fewer splits were 
required to isolate the sample.

\subparagraph{Likelihood.}
The negative sampling algorithm is implemented by converting the unsupervised 
outlier ranking problem into a semi-supervised 
problem~\citep{sipple:neg-sampling20}. Negative (anomalous) 
examples are created by sampling from an expanded space defined by the minimum 
and maximum values of each dimension of the positive (normal) examples. The 
negative and positive examples are then used to train a random forest 
classifier. We use the posterior probabilities of the random forest classifier 
as outlier scores, which means that the observations with higher posterior 
probabilities are more likely to be outliers. \todo{Describe PAE: Hannah or Bryce}
